### 2021-05-27 docs/checklist.tex
Updates from Overleaf

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerYes{}
  \item Did you describe the limitations of your work?
    \answerYes{Section \ref{section:conclusion} concludes that the implementation only covers co-dimension one bifurcations, with straightforward extensions to limit cycles, but does not scale well for partial differential equations}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerNA{We believe this work is sufficiently small scale and abstract to have any direct negative social impacts}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerYes{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerNA{In Section \ref{section:method} and the Appendix we apply existing theorems and approaches to differentiate our cost function}
	\item Did you include complete proofs of all theoretical results?
    \answerNA{While we do not have any theorems to prove, we include derivations in the Appendix to guide readers who are more theoretically inclined}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerYes{Figures can be reproduced using unit tests in the GitHub repository for \texttt{BifurcationFit.jl}}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerYes{In Section \ref{section:results} we details of optimisations with \texttt{Flux.jl}}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerYes{In Section \ref{section:results} we include error bars for time complexity estimates. We did not think that parameter uncertainties were relevant for the scope of this work}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerYes{The caption of Figure \ref{fig:scaling} has the specifications of the laptop used for complexity measurements}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerYes{\texttt{BifurcationKit.jl} \cite{Veltz2019PseudoArcLengthContinuation.jl} for continuation, \texttt{ForwardDiff.jl} \cite{Revels2016Forward-ModeJulia} for automatic differentiation and \texttt{Flux.jl} \cite{Innes2018FashionableFlux,Innes2018Flux:Julia} for optimisation}
  \item Did you mention the license of the assets?
    \answerNo{MIT Licences}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerYes{Julia package implementing our method \href{https://github.com/gszep/BifurcationFit.jl}{\texttt{github.com/gszep/BifurcationFit.jl}}}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerNA{No data was collected}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerNA{No data was collected}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerNA{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerNA{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerNA{}
\end{enumerate}

\end{enumerate}